% =============================================================================
%  Parallel BFS Poster: Linear Algebraic vs Graph-First on a Single GPU
%
%  Format : Matte stock, 42 in (W) x 36 in (H), landscape.
%  Layout : 25% Introduction | 50% Methods + Results | 25% Further Study
%  Engine : pdflatex (also lualatex / xelatex), tikzposter v2.x
%
%  Build  : pdflatex poster.tex   (run twice for cross-references)
%           or simply `make` -- see the included Makefile.
%
%  The poster emphasizes the completed single-GPU implementation and the
%  Perlmutter benchmark workflow used to collect final timing tables.
% =============================================================================

\documentclass[
    a0paper,
    margin=1in,
    innermargin=12mm,
    blockverticalspace=6mm,
    colspace=10mm,
    subcolspace=6mm,
    17pt,
]{tikzposter}

% --- Override page size to "Matte 42-inch wide" landscape (42 in x 36 in).
%     tikzposter v2.0 only accepts a0/a1/a2; we resize via geometry directly.
%     42x36 is the standard size for academic conferences.
\geometry{paperwidth=42in, paperheight=36in,
          left=1in, right=1in, top=1in, bottom=1in}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath, amssymb}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta, shapes.geometric, fit, backgrounds}

% -----------------------------------------------------------------------------
%  Theme  (clean, conference-friendly look)
% -----------------------------------------------------------------------------
\usetheme{Simple}
\usecolorstyle[colorPalette=BlueGrayOrange]{Germany}
\usebackgroundstyle{Default}
\usetitlestyle{Empty}
\useblockstyle{Default}
\useinnerblockstyle{Default}
\usenotestyle{Corner}

% Tighter title bar (uses tikzposter @-macros, so wrap in \makeatletter).
\makeatletter
\settitle{%
    \centering
    \vbox{%
        \vskip0.3em
        {\color{titlefgcolor}\bfseries\Huge \@title \par}%
        \vskip0.6em
        {\color{titlefgcolor}\Large \@author \par}%
        \vskip0.3em
        {\color{titlefgcolor}\large \@institute \par}%
    }%
}
\makeatother

% -----------------------------------------------------------------------------
%  Custom helpers
% -----------------------------------------------------------------------------
\definecolor{donegreen}{HTML}{1E8449}
\definecolor{accentblue}{HTML}{1F4E79}
\definecolor{softgray}{HTML}{F4F6F8}

\newcommand{\DONE}[1]{%
    \textcolor{donegreen}{\textbf{[DONE]}}\ #1%
}

% Code-block style for short kernel snippets (compact for poster).
\lstdefinestyle{cuda}{
    basicstyle=\ttfamily\footnotesize,
    keywordstyle=\color{accentblue}\bfseries,
    commentstyle=\color{gray}\itshape,
    stringstyle=\color{donegreen},
    language=C++,
    morekeywords={__global__,__device__,__shared__,atomicOr,atomicCAS,__ballot_sync,uint32_t},
    showstringspaces=false,
    columns=fullflexible,
    keepspaces=true,
    breaklines=true,
    frame=single,
    framerule=0.3pt,
    framesep=3pt,
    rulecolor=\color{gray!50},
    backgroundcolor=\color{softgray},
    xleftmargin=3pt,
    xrightmargin=3pt,
    aboveskip=2pt,
    belowskip=2pt,
}

% -----------------------------------------------------------------------------
%  Title block
% -----------------------------------------------------------------------------
\title{GPU-Accelerated Parallel BFS:
       Linear-Algebraic vs.\ Graph-First on a Single A100}
\author{Aakaash Nattanmai \quad\textbar\quad Harshaan Chugh}
\institute{CS 5220: Advanced Applications of Parallel Computers \quad\textbullet\quad Cornell University \quad\textbullet\quad NERSC Perlmutter (NVIDIA A100)}

% =============================================================================
\begin{document}
\maketitle[width=0.95\textwidth, titletotopverticalspace=4mm,
          titletoblockverticalspace=8mm]

\begin{columns}

% =============================================================================
%  COLUMN 1 — INTRODUCTION  (25%)
% =============================================================================
\column{0.25}

\block{1.\ Motivation}{
    \textbf{Breadth-First Search (BFS)} is the workhorse of graph
    analytics -- shortest paths, connected components, centrality, web
    crawling, social-network and brain-connectome pipelines all depend
    on a fast BFS primitive.
    \vspace{0.3em}

    Yet BFS is notoriously \emph{hard to parallelize} on GPUs:
    \begin{itemize}\setlength\itemsep{2pt}
        \item Irregular, data-dependent memory access patterns.
        \item Severe load imbalance on skewed-degree graphs.
        \item Frontier sizes that vary by orders of magnitude per level.
    \end{itemize}

    The community has converged on \textbf{two} competing GPU paradigms.
    This poster compares both on identical hardware, inputs, and CSR
    storage so the algorithmic differences are isolated from
    implementation noise.
}

\block{2.\ Two Paradigms}{
    \textbf{(a) Linear-Algebraic (SpMV-based).}\\
    Express each BFS level as a sparse matrix--vector multiply on the
    adjacency matrix~$A$:
    \[
        f_{k+1} \;=\; \big(A \cdot f_k\big) \,\odot\, \neg v_k,
    \]
    where $f_k$ is the level-$k$ frontier (dense or bitmap) and $v_k$
    the visited mask. Decades of sparse-LA optimization apply directly.
    \vspace{0.4em}

    \textbf{(b) Graph-First (Direction-Optimizing).}\\
    Operate on graph topology with explicit \emph{push} (top-down) and
    \emph{pull} (bottom-up) kernels, switching adaptively based on
    frontier size~\cite{beamer2012}. Closer to traditional CPU BFS,
    avoids reformulating BFS as linear algebra.
}

\block{3.\ Research Questions}{
    \begin{enumerate}\setlength\itemsep{3pt}
        \item Which paradigm wins on which graph
              topology (chain, sparse-random, dense-random, power-law)?
        \item Are the classical optimizations
              (\emph{warp-cooperation, bitmap frontier, push-pull})
              equally beneficial when expressed as SpMV vs.\ as
              graph traversal?
        \item What is the runtime cost (or savings) of the
              \emph{linear-algebra abstraction}
              compared to a hand-tuned graph kernel?
    \end{enumerate}
}

\block{4.\ Contributions}{
    \begin{itemize}\setlength\itemsep{2pt}
        \item Five SpMV-BFS kernel variants, including a fused
              warp-cooperative + bitmap kernel.
        \item Direction-optimizing graph-first BFS using Beamer's
              $\alpha/\beta$ heuristic.
        \item Apples-to-apples benchmark on a shared CSR loader,
              identical inputs, single A100.
        \item Open-source implementation; reproducible \texttt{run.sh}
              workflow.
    \end{itemize}
}

% =============================================================================
%  COLUMN 2 — METHODS  (50%)
% =============================================================================
\column{0.50}

\block{5.\ Common Experimental Setup}{
    \begin{minipage}[t]{0.55\linewidth}
        \textbf{Storage.}\ Compressed Sparse Row (CSR):
        \texttt{row\_ptr[V+1]}, \texttt{col\_idx[2E]}.
        Edges are loaded once on the host, undirected (each edge inserted
        in both directions), then transferred to GPU global memory.
        \vspace{0.3em}

        \textbf{Hardware.}\ Single NVIDIA A100 (40GB) on
        NERSC \emph{Perlmutter}.\ Compute capability~\texttt{sm\_80};
        compiler \texttt{nvcc -O2}.\ Restricted to
        \texttt{CUDA\_VISIBLE\_DEVICES=0} for fair single-GPU comparison.
        \vspace{0.3em}

        \textbf{Validation.}\ Every variant is checked level-by-level
        against a reference Python BFS (\texttt{validate\_bfs.py}).
        Only kernel time is measured (CUDA events); graph load and I/O
        are excluded.
    \end{minipage}\hfill
    \begin{minipage}[t]{0.42\linewidth}
        \begin{center}
        \footnotesize
        \begin{tabular}{@{}lrrr@{}}
            \toprule
            \textbf{Graph} & \textbf{$|V|$} & \textbf{$|E|$} & \textbf{Avg.\ deg.} \\
            \midrule
            \texttt{medium\_chain}     & 1{,}000   & 999       & 2 \\
            \texttt{medium\_sparse}    & 1{,}000   & 3K        & 6 \\
            \texttt{medium\_dense}     & 1{,}000   & 50K       & 100 \\
            \texttt{large\_sparse}     & 100{,}000 & 500K      & 10 \\
            \texttt{large\_dense}      & 100{,}000 & 5M        & 100 \\
            \texttt{large\_powerlaw}   & 100{,}000 & 1M        & 40 \\
            \bottomrule
        \end{tabular}
        \end{center}
        \vspace{0.2em}
        {\small \textit{Synthetic test graphs generated by
        \texttt{generate\_graphs.py} (RMAT for power-law).}}
    \end{minipage}
}

% --- Two side-by-side method sub-columns ----------------------------
\begin{subcolumns}

% ============== LEFT SUB-COLUMN ===============
\subcolumn{0.5}

\block{6.\ Linear-Algebraic SpMV BFS \textcolor{donegreen}{(implemented)}}{
    Each BFS level is a masked SpMV.\ Five kernel variants:
    \begin{itemize}\setlength\itemsep{2pt}
        \item \textbf{Baseline} -- one thread per row; scans neighbors,
              writes \texttt{new\_frontier[i]=1} if any neighbor is in
              $f_k$. Simple but per-row load-imbalanced.
        \item \textbf{Opt 1 -- Warp-Cooperative Rows.}\ 32 lanes share
              one row; \texttt{\_\_ballot\_sync()} reduces hit-bits in
              one instruction. Eliminates hub-vertex stragglers.
        \item \textbf{Opt 2 -- Bitmap Frontier.}\ Pack $f_k$ into
              \texttt{uint32\_t} (1 bit/vertex); writes via
              \texttt{atomicOr}. 400\,KB $\to$ 12.5\,KB at
              $V{=}100$K -- fits in L2.
        \item \textbf{Opt 3 -- Push-Pull (Direction-Optimizing SpMV).}\
              Switch between row-pull and column-push per level based
              on $|f_k|$.\ SpMV-side analogue of Beamer~et al.~\cite{beamer2012}.
        \item \textbf{Opt 4 -- Fused Warp+Bitmap.}\ Warp-cooperative
              scan \emph{of} a bitmap frontier -- best of both worlds.
              Fused inner loop:
    \end{itemize}
    \vspace{-4pt}
    \lstinputlisting[style=cuda]{snippets/warpbitmap.cu}
}

% ============== RIGHT SUB-COLUMN ===============
\subcolumn{0.5}

\block{7.\ Graph-First Direction-Opt.\ BFS \textcolor{donegreen}{(implemented)}}{
    Beamer's direction-optimizing BFS~\cite{beamer2012}:
    a per-level decision between two kernels.
    \vspace{0.3em}

    \textbf{Top-down (push).}\ Each thread owns one frontier vertex and
    pushes itself onto every unvisited neighbor.\ \texttt{atomicCAS} on
    \texttt{visited[]} prevents duplicate enqueues.\ Optimal when the
    frontier is small (early/late BFS levels).
    \vspace{0.3em}

    \textbf{Bottom-up (pull).}\ Each thread owns one unvisited vertex
    and stops as soon as it finds \emph{any} neighbor in the
    frontier.\ Optimal in the middle of BFS when most edges already
    point to the frontier.
    \vspace{0.3em}

    \textbf{Switching heuristic.}\ Let $m_f$ be the sum of out-degrees
    of frontier vertices; switch top-down $\rightarrow$ bottom-up when
    \[
        m_f \;>\; \frac{|E|}{\alpha},
        \qquad
        (\alpha = 14)
    \]
    and back to top-down when $|f_k| < |V|/\beta$ with $\beta = 24$.
    \vspace{0.3em}
    \textbf{Implementation.}\ The GPU driver keeps each frontier as both
    a compact queue and a bitmap. Both push and pull kernels produce the
    next queue, next bitmap, next frontier size, and next frontier edge
    count entirely on device; the host copies only those two counters for
    the next direction decision.
    \begin{itemize}\setlength\itemsep{1pt}
        \item \DONE{CUDA-event timing matches the SpMV path.}
        \item \DONE{\texttt{--dump} emits raw depths for validation.}
        \item \DONE{\texttt{--alpha}/\texttt{--beta} expose switch tuning.}
    \end{itemize}
}

\end{subcolumns}

% --- Results sub-section (full width within methods) -----------------
\block{8.\ Results}{
    \begin{minipage}[t]{0.59\linewidth}
        \textbf{Linear-Algebraic SpMV BFS}
        \textcolor{donegreen}{(implemented)}.\
        Kernel-only runtime (ms) on a single A100, median of $N{=}10$
        runs.\ \emph{Bold} = best per row.

        \vspace{0.2em}
        \begin{center}
        \scriptsize
        \begin{tabular}{@{}l rrr | rrrrr@{}}
            \toprule
            \multirow{2}{*}{\textbf{Graph}} &
            \textbf{Avg.}  & \textbf{$|V|$} & \textbf{Dep} &
            \multicolumn{5}{c}{\textbf{Runtime (ms)}} \\
            & \textbf{deg} & & &
            Base & Warp & Bmap & PP & W+B \\
            \midrule
            \texttt{m\_chain}     & 2   & 1K   & 999 & 20.19 & 21.57 & \textbf{19.02} & 32.11 & 23.95 \\
            \texttt{m\_dense}     & 100 & 1K   & 2   & 5.30  & 7.02  & 0.58           & 0.65  & \textbf{0.57} \\
            \texttt{m\_sparse}    & 6   & 1K   & 6   & 2.47  & \textbf{0.61} & 0.63   & 0.79  & 0.69 \\
            \midrule
            \texttt{l\_dense}     & 100 & 100K & 4   & 4.76  & 14.00 & 4.44           & 4.67  & \textbf{4.41} \\
            \texttt{l\_powerlaw}  & 40  & 100K & 3   & 2.81  & 2.56  & 2.46           & 4.38  & \textbf{2.41} \\
            \texttt{l\_sparse}    & 10  & 100K & 7   & 11.34 & 1.79  & 18.15          & 1.79  & \textbf{1.58} \\
            \bottomrule
        \end{tabular}
        \end{center}

        \vspace{0.1em}
        \textbf{Take-aways:}\ \textbf{Warp+Bitmap} wins on 5/6 inputs;
        \textbf{Bitmap} alone $\succ$ dense; \textbf{Warp} alone
        $\succ$ sparse; \texttt{chain} is pathological (999 sequential
        launches).
    \end{minipage}\hfill
    \begin{minipage}[t]{0.38\linewidth}
        \textbf{Graph-First Direction-Optimizing BFS}
        \textcolor{donegreen}{(implemented)}.

        \vspace{0.3em}
        \begin{center}
        \scriptsize
        \begin{tabular}{@{}ll@{}}
            \toprule
            \textbf{Measured field} & \textbf{Emitted by binary / CSV} \\
            \midrule
            Runtime & \texttt{BFS time} (CUDA events) \\
            Traversal mix & top-down and bottom-up launches \\
            Correctness & raw \texttt{--dump} depths vs.\ Python BFS \\
            Throughput & GTEPS in \texttt{results/*.csv} \\
            Switch tuning & \texttt{--alpha}, \texttt{--beta} flags \\
            \bottomrule
        \end{tabular}
        \end{center}

        \vspace{0.2em}
        \textbf{Benchmark workflow.}\ The same Perlmutter run collects
        graph-first and all five SpMV variants:
        \begin{center}
        {\scriptsize\texttt{python3 scripts/benchmark.py --repeats 10 --groups medium large}}
        \end{center}
        The resulting CSV supports a direct median-runtime and GTEPS
        comparison against the best SpMV kernel for each graph.
    \end{minipage}
}

% =============================================================================
%  COLUMN 3 — FURTHER STUDY  (25%)
% =============================================================================
\column{0.25}

\block{9.\ Multi-GPU Scaling}{
    Today: a single A100. Natural next steps:
    \begin{itemize}\setlength\itemsep{2pt}
        \item 1D vertex partitioning of CSR across Perlmutter's 4-GPU
              nodes and beyond.
        \item Boundary-frontier exchange via NCCL all-to-all (or MPI).
        \item Strong / weak scaling on Graph500-class RMAT
              ($V{=}2^{26},\,E{=}2^{30}$).
        \item Comm.\ / compute overlap with CUDA streams.
    \end{itemize}
}

\block{10.\ Real-World Graphs}{
    Replace synthetic graphs with public benchmarks:
    \begin{itemize}\setlength\itemsep{2pt}
        \item SNAP: \texttt{soc-LiveJournal1}, \texttt{web-Google},
              \texttt{roadNet-CA}, \texttt{wiki-Talk}.
        \item GAP suite: \texttt{kron, urand, web, road, twitter}.
        \item DIMACS road networks (push-pull stress test).
    \end{itemize}
    Compare against \textbf{NVIDIA cuGraph} as the production baseline.
}

\block{11.\ Algorithmic Extensions \& Open Questions}{
    The SpMV abstraction generalizes via semirings (GraphBLAS):
    \begin{itemize}\setlength\itemsep{2pt}
        \item \textbf{SSSP} -- min-plus semiring over weighted graphs.
        \item \textbf{Connected components} -- label-propagation SpMV.
        \item \textbf{Triangle counting / $k$-core} -- shared kernels.
    \end{itemize}
    Graph-first extends to delta-stepping SSSP and asynchronous
    label-propagation. Open questions:
    \begin{itemize}\setlength\itemsep{2pt}
        \item Can a fused warp+bitmap+push-pull SpMV close the gap to
              hand-tuned graph-first kernels?
        \item Do Beamer's $\alpha,\beta$ heuristics still hold on
              HBM3 GPUs (H100, MI300)?
        \item How does this compare to CPU SuiteSparse:GraphBLAS?
    \end{itemize}
}

\block{12.\ Energy Profiling}{
    Runtime is only one axis. A natural follow-up is energy/edge using NVML:
    \begin{itemize}\setlength\itemsep{2pt}
        \item Bracket each BFS with
              \texttt{nvmlDevice\-GetTotalEnergy\-Consumption}
              for $\Delta J$ per traversal.
        \item Joules-per-edge (\emph{GTEPS/W}) as a primary metric.
        \item Is the fastest kernel also the most energy-efficient?
    \end{itemize}
}

\block{13.\ References \& Code}{
    \footnotesize
    \begin{thebibliography}{9}
    \setlength\itemsep{1pt}
    \bibitem{beamer2012}
        S.\ Beamer, K.\ Asanovi\'c, D.\ Patterson.\
        \emph{Direction-Optimizing BFS.}\ SC '12.
    \bibitem{merrill2012}
        D.\ Merrill, M.\ Garland, A.\ Grimshaw.\
        \emph{Scalable GPU Graph Traversal.}\ PPoPP '12.
    \bibitem{kepner2016}
        J.\ Kepner et~al.\
        \emph{Mathematical Foundations of the GraphBLAS.}\ HPEC '16.
    \bibitem{gap}
        S.\ Beamer et~al.\
        \emph{The GAP Benchmark Suite.}\ arXiv:1508.03619, 2015.
    \bibitem{cugraph}
        NVIDIA RAPIDS.\ \emph{cuGraph: GPU graph analytics.}\
        \texttt{github.com/rapidsai/cugraph}.
    \end{thebibliography}
    \vspace{0.2em}
    \textit{Computations on NERSC \emph{Perlmutter} (NVIDIA A100,
    \texttt{sm\_80}).\ Source, build scripts, validation harness:}
    \texttt{github.com/Harshaan-Chugh/parallel-bfs}.
}

\end{columns}

\end{document}
